### 2017-10-17 paper/sections/1_introduction.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is the problem?
File system metadata services in HPC have scalability problems. It has been
shown that HPC workloads are metadata resource intensive because administrative
tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
system~\cite{zheng:pdsw2014-batchfs}, on large data sets lead to contention
for the same directories and inodes ({\it
e.g.}, path traversal). Applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every client is unreasonable. This problem is exacerbated
by current trends in HPC, where architectures are transitioning from complex
storage stacks with burst buffer, file system, object store, and tape tiers to
more simplified stacks with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
because more requests end up hitting the same layer and latencies cannot be 
hidden while data migrates across tiers.

\begin{figure}[tb]
\centering
\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
\caption{With CudeleFS, users can assign weaker consistency and durability
policies to subtrees to get the same performance benefits of state-of-the-art
HPC architectures, such as those shown in the embedded table, while
applications that require the stronger guarantees of POSIX can still reside in
the same namespace.
}\label{fig:subtree-policies}
\end{figure}

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, batch style jobs often do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
patterns).

@@

% What did we do
To address this problem, we propose a subtree policy API that lets developers dynamically control the
consistency and durability guarantees for subtrees in the file system
namespace. For example, one subtree can adopt weaker consistency semantics
while another subtree can retain the rigidity of POSIX's strong consistency.
The subtrees in Figure~\ref{fig:subtree-policies} show a setup in our proposed system where a single global
namespace has subtrees for applications optimized with techniques from different
state-of-the-art HPC architectures.

In this work, we present CudeleFS, a prototype programmable file system that
supports different degrees of consistency and durability in a global namespace.
CudeleFS achieves this by exposing ``mechanisms" that developers use to specify
their preferred semantics.  CudeleFS supports 3 forms of consistency
(invisible, weak, and strong) and 3 degrees of durability (none, local, and
global) giving the user a wide range of policies and optimizations that can be
custom fit to an application. We make the following contributions:

@@

  \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
  previously unexplored metadata designs, all implemented using CudeleFS.

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.
In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the CudeleFS
prototype and API. Section~\ref{sec:implementation} describes the CudeleFS
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places CudeleFS in the context of other related
work.

>>>>

% What is the problem?
File system metadata services in HPC have scalability problems. Administrative
tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
system~\cite{zheng:pdsw2014-batchfs} lead to contention for the same
directories and inodes ({\it e.g.}, path traversal). Applications perform
better with dedicated metadata servers~\cite{sevilla:sc15-mantle,
ren:sc2014-indexfs} but provisioning a metadata server for every client is
unreasonable. This problem is exacerbated by current trends in HPC, where
architectures are transitioning from complex storage stacks with burst buffer,
file system, object store, and tape tiers to more simplified stacks with just a
burst buffer and object store~\cite{bent:login16-hpc-trends}; this puts more
pressure on data access because more requests end up hitting the same layer and
latencies cannot be hidden while data migrates across tiers.

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
\caption{ Clients that need better performance can relax consistency on their
subtree ({e.g.}, HDFS) or even decouple the subtree and move it locally
({e.g.}, BatchFS, RAMDisk). Decoupled subtrees can further relax durability
semantics to achieve even better performance. Meanwhile, clients that require
stronger guarantees ({\it e.g.,} POSIX) can still reside in the same namespace.
}\label{fig:subtree-policies}
\end{figure}

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, many batch style jobs do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
patterns).

@@

% What did we do
To address this problem, we present an API and framework that lets developers
dynamically control the consistency and durability guarantees for subtrees in
the file system namespace.  Figure~\ref{fig:subtree-policies} shows a setup in
our proposed system where a single global namespace has subtrees for
applications optimized with techniques from different state-of-the-art HPC
architectures.  The HDFS\footnote{This is example is not directly evaluated in
this paper, although a similiar use case is explored in~\S\ref{use-case-1}}
subtree has weaker consistency than strong consistency because it  does not
guarantee that all updates are not immediately seen by all servers; the BatchFS
and RAMDisk subtrees are decoupled from the global namespace and have similiar
consistency/durability behavior to those systems; and the POSIX subtree retains
the rigidity of POSIX's strong consistency.

Our protoype system, Cudele, achieves this by exposing ``mechanisms" that
developers use to specify their preferred semantics.  Cudele supports 3 forms
of consistency (invisible, weak, and strong) and 3 degrees of durability (none,
local, and global) giving the user a wide range of policies and optimizations
that can be custom fit to an application. We make the following contributions:

@@

  \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
  previously unexplored metadata designs, all implemented using Cudele.

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures. Because
this paper focuses on file system metadata protocols, we only scale up to 20
clients executing an extremely metdata intensive workload; this load is the
maximum capacity of a single metadata server for the platform we chose to
implement Cudele on.  Partitioning and replicating metadata across a cluster
can be controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in
this paper we show the capacity and performance of a metadata server with
superior metadata protocols, which should be used to inform load balancing
decisions.

In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
cost of POSIX consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
prototype and API. Section~\ref{sec:implementation} describes the Cudele
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places Cudele in the context of other related
work.

### 2017-10-17 paper/sections/4_methodology.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications decouple the namespace, write updates to a local journal,
and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
\end{figure}

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{The CudeleFS mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

In this section we describe CudeleFS, our file system that lets users
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. CudeleFS exposes these
mechanisms and the user composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

\subsection{The CudeleFS Mechanisms}
\label{sec:the-cudelesfs-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Append Client Journal"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisfied by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
client and applies the updates directly to the in-memory metadata store maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- CudeleFS makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

@@

%The metadata store and journal
%are different ways of representing the namespace.  CudeleFS presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Append Client Journal" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

@@

\subsection{Defining Policies in CudeleFS}
\label{sec:setting-policies-with-cudele}

@@

\caption{Users can explore the consistency (C) and
durability (D) spectrums by composing CudeleFS mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

@@

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed to have ``global" durability until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that changes a subtree from ``local
persist" to ``global persist", CudeleFS makes no guarantee until the mechanisms are
complete.

\subsection{CudeleFS Namespace API}
\label{sec:cudelefs-namespace-api}

@@

The ``Consistency" and ``Durability" parameters are set to the CudeleFS
mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
``Allocated Inodes" is a way for the application to specify how many files it
intends to create. It is a contract so that the file system can provision
enough resources for the incumbent merge and so it can give valid inodes to
other clients. The inodes can be used anywhere within the decoupled namespace
({i.e.} at any depth in the subtree).

@@

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
would use the configuration:

>>>>

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications decouple the namespace, write updates to a local journal,
and delay metadata updates using the Cudele Mechanisms }\label{fig:decouple}
\end{figure}

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{The Cudele mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

In this section we describe Cudele, our file system that lets users
assign consistency and durability semantics to subtrees in the global
namespace. A \textbf{mechanism} is an abstraction and basic building block for
constructing consistency and durability guarantees. Cudele exposes these
mechanisms and the user composes them together to construct
\textbf{policies}. These policies are assigned to subtrees and they dictate how
the file system should handle operations within that subtree.  Below, we
describe the mechanisms, the policies, and the API for assigning policies to
subtrees.

\subsection{The Cudele Mechanisms}
\label{sec:the-cudelesfs-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Append Client Journal"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls Cudele mechanisms to achieve the desired
consistency/durability semantics.  Cudele provides a library for clients to
link into and all operations are performed by the client.

\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
calls for every metadata operation from the client to the metadata server,
assuming the request cannot be satisfied by the inode cache. This mechanism is
part of the default CephFS implementation and is the strongest form of
consistency because clients see metadata updates right away.  ``Nonvolatile
Apply" replays the client's journal onto the metadata cluster's metadata store.
The client's in-memory journal is written into the object store and the
metadata servers are restarted. When the metadata servers re-initialize, they
notice new journal updates in the object store and replay the events onto their
in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
client and applies the updates directly to the in-memory metadata store maintained
by the metadata servers. We say volatile because -- in exchange for peak
performance -- Cudele makes no consistency or durability guarantees while
``Volatile Apply" is executing.  If a concurrent update from a client occurs
there is no rule for resolving conflicts and if the client or metadata server
crashes there may be no way to recover.

@@

%The metadata store and journal
%are different ways of representing the namespace.  Cudele presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Append Client Journal" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

@@

\subsection{Defining Policies in Cudele}
\label{sec:setting-policies-with-cudele}

@@

\caption{Users can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. 
\label{table:spectrum}}
\end{center}
\end{table}

@@

% talk of eventual consistency
The consistency and durability properties in Table~\ref{table:spectrum} are not
guaranteed until all mechanisms in the cell are complete. The compositions
should be considered atomic and there are no guarantees while transitioning
between policies. For example, updates are not deemed to have ``global" durability until
they are safely saved in the object store. If a failure occurs during ``global
persist" or if we inject a new policy that changes a subtree from ``local
persist" to ``global persist", Cudele makes no guarantee until the mechanisms are
complete.

\subsection{Cudele Namespace API}
\label{sec:cudelefs-namespace-api}

@@

The ``Consistency" and ``Durability" parameters are set to the Cudele
mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
``Allocated Inodes" is a way for the application to specify how many files it
intends to create. It is a contract so that the file system can provision
enough resources for the incumbent merge and so it can give valid inodes to
other clients. The inodes can be used anywhere within the decoupled namespace
({i.e.} at any depth in the subtree).

@@

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on Cudele, the user
would use the configuration:

### 2017-10-17 paper/sections/5_implementation.tex
paper: re-org eval and intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
CudeleFS; namely, we try to leverage components inside Ceph to inherit the
robustness and correctness of the internal subsystems. Using this `dirty-slate'
approach, we only had to implement 4 of the 6 mechanisms from
Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
system itself.  In this section, we first describe a Ceph internal subsystem or
component and then we show how we use it in CudeleFS.

@@

% how i used it
\noindent\textbf{CudeleFS}: the ``RPCs" mechanism uses the metadata store to
service requests. Using code designed for recovery, ``Nonvolatile Apply" and
``Volatile Apply" replay updates onto the metadata store in memory and in the
object store, respectively.  When the clients are ready to merge their updates
back into the global namespace, they pass a binary file of  metadata updates to
the metadata server.

@@

\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Append Client Journal",
``Local Persist", and ``Global Persist".  ``Stream" is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemented by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, CudeleFS leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

@@

\noindent\textbf{CudeleFS}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
``Append Client Journal", ``Volatile Apply", and ``Nonvolatile Apply".

@@

\noindent\textbf{CudeleFS}: ``Nonvolatile Apply" uses the internal inode cache
code to allocate inodes to clients that decouple parts of the namespace and to
skip inodes used by the client at merge time.\\

@@

\noindent\textbf{CudeleFS}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different CudeleFS mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.

>>>>

We use a programmable storage approach~\cite{sevilla:eurosys17} to design
Cudele; namely, we try to leverage components inside Ceph to inherit the
robustness and correctness of the internal subsystems. Using this `dirty-slate'
approach, we only had to implement 4 of the 6 mechanisms from
Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
system itself.  In this section, we first describe a Ceph internal subsystem or
component and then we show how we use it in Cudele.

@@

% how i used it
\noindent\textbf{Cudele}: the ``RPCs" mechanism uses the metadata store to
service requests. Using code designed for recovery, ``Nonvolatile Apply" and
``Volatile Apply" replay updates onto the metadata store in memory and in the
object store, respectively.  When the clients are ready to merge their updates
back into the global namespace, they pass a binary file of  metadata updates to
the metadata server.

@@

\noindent\textbf{Cudele}: the journal format is used by ``Stream", ``Append Client Journal",
``Local Persist", and ``Global Persist".  ``Stream" is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemented by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, Cudele leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

@@

\noindent\textbf{Cudele}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
``Append Client Journal", ``Volatile Apply", and ``Nonvolatile Apply".

@@

\noindent\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
code to allocate inodes to clients that decouple parts of the namespace and to
skip inodes used by the client at merge time.\\

@@

\noindent\textbf{Cudele}: uses the large inodes to store consistency and
durability policies in the directory inode. This approach uses the File Type
interface from the Malacology programmable store
system~\cite{sevilla:eurosys17} and it tells clients how to access the
underlying metadata. The underlying implementation stores executable code in
the inode that calls the different Cudele mechanisms. Of course, there are many
security and access control aspects of this approach but that is beyond the
scope of this paper.
